---
title: "PredictiveFlims - The Movie Recommendation System"
author: "Durvesh Chaudhari(AI-A-02) , Mayank Kulkarni(AI-A-11) , Sanket Kulkarni(AI-A-12), Rushikesh Tanpure(AI-B-08) , Ganesh Thorat(AI-B-09) "
output:
  word_document: default
  html_notebook: default
  pdf_document: default
---
Importing Library
```{r}
library(recommenderlab)     #for recommendation
library(reshape2)
library(data.table)
library(ggplot2)

```


#Importing the DataSet
```{r}
movie_data<-read.csv("D:/BTech/2nd_Year/Data Science/DS_Project/movies.csv",stringsAsFactors = FALSE)
rating_data<-read.csv("D:/BTech/2nd_Year/Data Science/DS_Project/ratings.csv")
head(movie_data)
```
```{r}
head(rating_data)
```

Structure of data
```{r}
str(movie_data)
str(rating_data)
```
#tabular View
```{r}
data.table(movie_data)
```
#tabular View
```{r}
data.table(rating_data)
```

Summarizing the data
```{r}
summary(movie_data)
summary(rating_data)
```

We need to add more information with genre so we will create another dataframe
```{r}
movie_genre<- as.data.frame(movie_data$genres,stringsAsFactors = FALSE)
head(movie_genre)
movie_genre2<-as.data.frame(tstrsplit(movie_genre[,1],"[|]",type.convert = TRUE),stringsAsFactors = FALSE)
head(movie_genre2)
```
```{r}
colnames(movie_genre2)<-c(1:10)
list_genre<-c("Action","Adventure","Animation","Children","Comedy","Crime","Documentary","Drama","Fantasy","Film-Noir","Horror","Musical","Mystery","Romance","Sci-Fi","Thriller","war","western")
genre_matl <- matrix(0,10330,18)
genre_matl[1,]<-list_genre
colnames(genre_matl) <-list_genre

for(index in 1:nrow(movie_genre2))
{
  for(col in 1:ncol(movie_genre2))
  {
    gen_col=which(genre_matl[1,]==movie_genre2[index,col])
    genre_matl[index+1,gen_col]<-1
  }
}

print("First genre_matl operations are completed")
```

```{r}
genre_mat2<-as.data.frame(genre_matl[-1,],stringsAsFactors = FALSE)

for(col in 1:ncol(genre_mat2))
{
  genre_mat2[,col]<-as.integer(genre_mat2[,col])
}

str(genre_mat2)
```

create search matrix that gives us films based on genres
```{r}
SearchMovie <- cbind(movie_data[,1:2],genre_mat2[])
head(SearchMovie)
```

Many Movies have several genre... So Let's create sparse matrix for recommendation
```{r}
#it is a spare matrix
ratingMatrix <-dcast(rating_data,userId~movieId,value.var = "rating",na.rm=FALSE)    

ratingMatrix<-as.matrix(ratingMatrix[,-1])         #Don't need user id so removeing it
ratingMatrix<-as(ratingMatrix,"realRatingMatrix")  #Converting rating matrix to spare matrix
head(ratingMatrix)
```

!!!!!!!-----------------Recommendation Model-------------------------!!!!!!!


```{r}
recommendation_model<-recommenderRegistry$get_entries(dataType="realRatingMatrix")
names(recommendation_model)
```

```{r}
lapply(recommendation_model,"[[" ,"description")

```
we will use item based collaborative filtering
```{r}
recommendation_model$IBCF_realRatingMatrix$parameters
```

Let's check similarity 
```{r}
similarity_mat<-similarity(ratingMatrix[1:4, ],method = "cosine",which="users")
as.matrix(similarity_mat)
image(as.matrix(similarity_mat),main="User's Similarity")
```
Let's check similarity of movies
```{r}
movie_similarity<-similarity(ratingMatrix[,1:4],method = "cosine",which="items")
as.matrix(movie_similarity)
image(as.matrix(movie_similarity),main="Moive Similarity")
```

Rating Values
```{r}
rating_values<-as.vector(ratingMatrix@data)
unique(rating_values)
```

how much rating as count of numbers
```{r}
Table_rating<-table(rating_values)
Table_rating
```

Most viewed movies visualiztion
```{r}
library(ggplot2)
movie_views<-colCounts(ratingMatrix)    #count the views for each movie
table_views<-data.frame(movie=names(movie_views),views=movie_views) #create dataframe of views
table_views<-table_views[order(table_views$views,decreasing=TRUE),]  #sort number of views
table_views$title<-NA

for (index in 1:10325)
{
    table_views[index,3]<-as.character(subset(movie_data,movie_data$movieId==table_views[index,1])$title)
}

table_views[1:6,]
```
# Visualize a bar plot for the total number of views of the top films
```{r}
ggplot(table_views[1:6, ], aes(x = title, y = views)) +
  geom_bar(stat="identity", fill = 'steelblue') +
  geom_text(aes(label=views), vjust=-0.3, size=3.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("Total Views of the Top Films")
```
HeatMap of rating
```{r}
image(ratingMatrix[1:30,1:30],axes=FALSE,main="30 x 30 hatmap")
```
lot's of sparse data
Now we will perform below operation
1}Select Useful data 
2}Normalize Data
3}Binarize Data

We have seen the rating datset. So now what do you think how many users need to rate a movie
```{r}
movie_ratings <- ratingMatrix[rowCounts(ratingMatrix) > 50,
                              colCounts(ratingMatrix) > 50]
movie_ratings
```
describing matrix of relevant users
```{r}
minimum_movies<- quantile(rowCounts(movie_ratings), 0.98)
minimum_users <- quantile(colCounts(movie_ratings), 0.98)
image(movie_ratings[rowCounts(movie_ratings) > minimum_movies,
                    colCounts(movie_ratings) > minimum_users],
      main = "Heatmap of the top users and movies")
```
Visualizing the distribution of the average ratings per user
```{r}
average_ratings <- rowMeans(movie_ratings)
qplot(average_ratings, fill=I("steelblue"), col=I("red")) +
  ggtitle("Distribution of the average rating per user")
```

2}Data Normalization
```{r}
normalized_ratings <- normalize(movie_ratings)
sum(rowMeans(normalized_ratings) > 0.00001)
image(normalized_ratings[rowCounts(normalized_ratings) > minimum_movies,
                         colCounts(normalized_ratings) > minimum_users],
      main = "Normalized Ratings of the Top Users")
```
3} Data Binarization (Binary means 0 and 1 will be recommended if rating of that movie is greater than 3.5)
```{r}
binary_minimum_movies <- quantile(rowCounts(movie_ratings), 0.90)
binary_minimum_users <- quantile(colCounts(movie_ratings), 0.90)
movies_watched <- binarize(movie_ratings, minRating = 1)
good_rated_films <- binarize(movie_ratings, minRating = 3.5)
image(good_rated_films[rowCounts(movie_ratings) > binary_minimum_movies,
                       colCounts(movie_ratings) > binary_minimum_users],
      main = "Heatmap of the top users and movies")
```
# Collaborative Filtering System
# Splitting the dataset into 80% training set and 20% test set
```{r}
sampled_data<- sample(x = c(TRUE, FALSE),
                      size = nrow(movie_ratings),
                      replace = TRUE,
                      prob = c(0.8, 0.2))
training_data <- movie_ratings[sampled_data, ]
testing_data <- movie_ratings[!sampled_data, ]
```
# Building the Recommendation System
```{r}
recommendation_system <- recommenderRegistry$get_entries(dataType ="realRatingMatrix")
recommendation_system$IBCF_realRatingMatrix$parameters
```
```{r}
recommen_model <- Recommender(data = training_data,
                              method = "IBCF",
                              parameter = list(k = 30))
recommen_model
class(recommen_model)
```
# Exploring the data science recommendation system model
```{r}
model_info <- getModel(recommen_model)

class(model_info$sim)
dim(model_info$sim)
top_items <- 20
image(model_info$sim[1:top_items, 1:top_items],
      main = "Heatmap of the first rows and columns")
```
# Visualize sum of rows and columns with the similarity of the objects above 0
```{r}
sum_rows <- rowSums(model_info$sim > 0)
table(sum_rows)
sum_cols <- colSums(model_info$sim > 0)
qplot(sum_cols, fill=I("steelblue"), col=I("red"))+ ggtitle("Distribution of the column count")
```

Let's Recommend the Movie....
# the number of items to recommend to each user
```{r}
top_recommendations <- 10 
predicted_recommendations <- predict(object = recommen_model,
                                     newdata = testing_data,
                                     n = top_recommendations)
predicted_recommendations
```

Let's See the some of names
# recommendation for the first user
```{r}
user1 <- predicted_recommendations@items[[1]] 
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie_data,
                                             movie_data$movieId == movies_user1[index])$title)
}
movies_user2
```
```{r}
user1 <- predicted_recommendations@items[[2]] 
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie_data,
                                             movie_data$movieId == movies_user1[index])$title)
}
movies_user2

```
```{r}

user1 <- predicted_recommendations@items[[3]] 
movies_user1 <- predicted_recommendations@itemLabels[user1]
movies_user2 <- movies_user1
for (index in 1:10){
  movies_user2[index] <- as.character(subset(movie_data,
                                             movie_data$movieId == movies_user1[index])$title)
}
movies_user2
```




